---
title: "chapter 1: R Basics"
output: html_notebook
---
<!-- # Chapter 1. R Basics -->

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [1.1. Installing a Package](#11-installing-a-package)
- [1.2. Loading a Package](#12-loading-a-package)
- [1.3. Loading a Delimited Text Data File](#13-loading-a-delimited-text-data-file)
- [1.4. Loading Data from an Excel File](#14-loading-data-from-an-excel-file)
- [1.5. Loading Data from an SPSS File](#15-loading-data-from-an-spss-file)

<!-- tocstop -->

## 1.1. Installing a Package
  * Problem  
    You want to install a package from CRAN.  

  * Solution  
```{r}  
    install.packages("ggplot2" )  
```
  * Discussion  

## 1.2. Loading a Package
  * Problem  
    You want to load an installed package.  

  * Solution ```{R}
library(ggplot2```  
  * Discussion  

## 1.3. Loading a Delimited Text Data File
  * Problem  
    You want to load data from a delimited text file.  

  * Solution  
```{r}
    data <-  read.csv("datafile.csv" )
```
  * Discussion  
```{R}
# ====================================
data <- read.csv("datafile.csv", header=FALSE)

# ====================================
# Manually assign the header names
names(data) <- c("Column1","Column2","Column3")

# ====================================
data <- read.csv("datafile.csv", sep="\t")

# ====================================
data <- read.csv("datafile.csv", stringsAsFactors=FALSE)
# Convert to factor
data$Sex <- factor(data$Sex)
str(data)
```  
  * See Also  

## 1.4. Loading Data from an Excel File
  * Problem  
    You want to load data from an Excel file.  

  * Solution  
```{r}
# Only need to install once
install.packages("xlsx")

library(xslx)
data <- read.xlsx("datafile.xlsx", 1)
```    
```{r}
# Only need to install once
install.packages("gdata")

library(gdata)
# Read first sheet
data <- read.xls("datafile.xls")
```  

  * Discussion  
```{R}
# ====================================
data <- read.xlsx("datafile.xls", sheetIndex=2)

data <- read.xlsx("datafile.xls", sheetName="Revenues")


# ====================================
data <- read.xls("datafile.xls", sheet=2)
```


  * See Also  

## 1.5. Loading Data from an SPSS File
  * Problem  
    You want to load data from an SPSS file.  

  * Solution  
```{r}  
# Only need to install the first time
install.packages("foreign")

library(foreign)
data <- read.spss("datafile.sav")  
```
  * Discussion  
    The foreign package also includes functions to load from other formats, including:
    * read.octave(): Octave and MATLAB
    * read.systat(): SYSTAT
    * read.xport(): SAS XPORT
    * read.dta(): Stata
  * See Also  
