---
title: 'chapter2: Quickly Exploring Data'
output: html_notebook
---

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [2.1. Creating a Scatter Plot](#21-creating-a-scatter-plot)
- [2.2. Creating a Line Graph](#22-creating-a-line-graph)
- [2.3. Creating a Bar Graph](#23-creating-a-bar-graph)
- [2.4. Creating a Histogram](#24-creating-a-histogram)
- [2.5. Creating a Box Plot](#25-creating-a-box-plot)
- [2.6. Plotting a Function Curve](#26-plotting-a-function-curve)

<!-- tocstop -->

## 2.1. Creating a Scatter Plot
  * Problem  
    You want to create a scatter plot.  

  * Solution  
```{r}  
plot(mtcars$wt, mtcars$mpg)
```


```{r}
library(ggplot2)
qplot(mtcars$wt, mtcars$mpg)
```

```{r}    
qplot(wt, mpg, data=mtcars)
# This is equivalent to:
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()
```  
  * See Also  

## 2.2. Creating a Line Graph
  * Problem  
    You want to create a line graph.  

  * Solution  
```{r}
# ====================================
plot(pressure$temperature, pressure$pressure, type="l")


# ====================================
plot(pressure$temperature, pressure$pressure, type="l")
points(pressure$temperature, pressure$pressure)

lines(pressure$temperature, pressure$pressure/2, col="red")
points(pressure$temperature, pressure$pressure/2, col="red")


# ====================================
library(ggplot2)
qplot(pressure$temperature, pressure$pressure, geom="line")

# ====================================
qplot(temperature, pressure, data=pressure, geom="line")

# This is equivalent to:
ggplot(pressure, aes(x=temperature, y=pressure)) + geom_line()

# Lines and points together
qplot(temperature, pressure, data=pressure, geom=c("line", "point"))

# Equivalent to:
ggplot(pressure, aes(x=temperature, y=pressure)) + geom_line() + geom_point()

```
  * See Also  

## 2.3. Creating a Bar Graph
  * Problem  
    You want to make a bar graph.  

  * Solution  
```{r}
# ====================================
barplot(BOD$demand, names.arg=BOD$Time)


# ====================================
table(mtcars$cyl)

 4  6  8
11  7 14
# There are 11 cases of the value 4, 7 cases of 6, and 14 cases of 8



# ====================================
# Generate a table of counts
barplot(table(mtcars$cyl))


# ====================================
library(ggplot2)
qplot(BOD$Time, BOD$demand, geom="bar", stat="identity")

# Convert the x variable to a factor, so that it is treated as discrete
qplot(factor(BOD$Time), BOD$demand, geom="bar", stat="identity")


# ====================================
# cyl is continuous here
qplot(mtcars$cyl)

# Treat cyl as discrete
qplot(factor(mtcars$cyl))


# ====================================
# Bar graph of values. This uses the BOD data frame, with the
#"Time" column for x values and the "demand" column for y values.
qplot(Time, demand, data=BOD, geom="bar", stat="identity")
# This is equivalent to:
ggplot(BOD, aes(x=Time, y=demand)) + geom_bar(stat="identity")


# Bar graph of counts
qplot(factor(cyl), data=mtcars)
# This is equivalent to:
ggplot(mtcars, aes(x=factor(cyl))) + geom_bar()
```
  * See Also  

## 2.4. Creating a Histogram
  * Problem  
    You want to view the distribution of one-dimensional data with a histogram.  

  * Solution  
```{r}
# ====================================
hist(mtcars$mpg)

# Specify approximate number of bins with breaks
hist(mtcars$mpg, breaks=10)


# ====================================
qplot(mtcars$mpg)


# ====================================
library(ggplot2)
qplot(mpg, data=mtcars, binwidth=4)
# This is equivalent to:
ggplot(mtcars, aes(x=mpg)) + geom_histogram(binwidth=4)
```
  * See Also  

## 2.5. Creating a Box Plot
  * Problem  

  * Solution  

  * See Also  

## 2.6. Plotting a Function Curve
  * Problem  

  * Solution  

  * See Also  
