---
title: "Chapter 7. Annotations"
output: html_notebook
---

# Chapter 7. Annotations

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [Chapter 7. Annotations](#chapter-7-annotations)
	- [7.1. Adding Text Annotations](#71-adding-text-annotations)
	- [7.2. Using Mathematical Expressions in Annotations](#72-using-mathematical-expressions-in-annotations)
	- [7.3. Adding Lines](#73-adding-lines)
	- [7.4. Adding Line Segments and Arrows](#74-adding-line-segments-and-arrows)
	- [7.5. Adding a Shaded Rectangle](#75-adding-a-shaded-rectangle)
	- [7.6. Highlighting an Item](#76-highlighting-an-item)
	- [7.7. Adding Error Bars](#77-adding-error-bars)
	- [7.8. Adding Annotations to Individual Facets](#78-adding-annotations-to-individual-facets)

<!-- tocstop -->

## 7.1. Adding Text Annotations
* **Problem**  
  You want to add a text annotation to a plot.  

* **Solution**  
```{R}
p <- ggplot(faithful, aes(x=eruptions, y=waiting)) + geom_point()

p + annotate("text", x=3, y=48, label="Group 1") +
    annotate("text", x=4.5, y=66, label="Group 2")
```

* **Discussion**  
```{r}
# ====================================
p + annotate("text", x=3, y=48, label="Group 1", family="serif",
             fontface="italic", colour="darkred", size=3) +
    annotate("text", x=4.5, y=66, label="Group 2", family="serif",
             fontface="italic", colour="darkred", size=3)


# ====================================
p + annotate("text", x=3, y=48, label="Group 1", alpha=.1) +     # Normal
    geom_text(x=4.5, y=66, label="Group 2", alpha=.1)            # Overplotted


# ====================================
p + annotate("text", x=-Inf, y=Inf, label="Upper left", hjust=-.2, vjust=2) +
    annotate("text", x=mean(range(faithful$eruptions)), y=-Inf, vjust=-0.4,
             label="Bottom middle")

```    

* **See Also**  

## 7.2. Using Mathematical Expressions in Annotations
* **Problem**  
  You want to add a text annotation with mathematical notation.  

* **Solution**  
```{R}
# A normal curve
p <- ggplot(data.frame(x=c(-3,3)), aes(x=x)) + stat_function(fun = dnorm)

p + annotate("text", x=2, y=0.3, parse=TRUE,
             label="frac(1, sqrt(2 * pi)) * e ^ {-x^2 / 2}")
```

* **Discussion**  
```{r}
p + annotate("text", x=0, y=0.05, parse=TRUE, size=4,
             label="'Function:  ' * y==frac(1, sqrt(2*pi)) * e^{-x^2/2}")
```             

* **See Also**  

## 7.3. Adding Lines
* **Problem**  
  You want to add lines to a plot.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

p <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) + geom_point()

# Add horizontal and vertical lines
p + geom_hline(yintercept=60) + geom_vline(xintercept=14)

# Add angled line
p + geom_abline(intercept=37.4, slope=1.75)

```

* **Discussion**  
```{R}
# ====================================
library(plyr) # For the ddply() function
hw_means <- ddply(heightweight, "sex", summarise, heightIn=mean(heightIn))
hw_means

 sex heightIn
   f 60.52613
   m 62.06000

p + geom_hline(aes(yintercept=heightIn, colour=sex), data=hw_means,
               linetype="dashed", size=1)



# ====================================
pg <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_point()

pg + geom_vline(xintercept = 2)

pg + geom_vline(xintercept = which(levels(PlantGrowth$group)=="ctrl"))
```

* **See Also**  

## 7.4. Adding Line Segments and Arrows
* **Problem**  
  You want to add line segments or arrows to a plot.  

* **Solution**  
```{r}
library(gcookbook) # For the data set

p <- ggplot(subset(climate, Source=="Berkeley"), aes(x=Year, y=Anomaly10y)) +
     geom_line()

p + annotate("segment", x=1950, xend=1980, y=-.25, yend=-.25)

```

* **Discussion**  
```{R}
library(grid)
p + annotate("segment", x=1850, xend=1820, y=-.8, yend=-.95, colour="blue",
             size=2, arrow=arrow()) +
    annotate("segment", x=1950, xend=1980, y=-.25, yend=-.25,
             arrow=arrow(ends="both", angle=90, length=unit(.2,"cm")))
```             

* **See Also**  

## 7.5. Adding a Shaded Rectangle
* **Problem**  
  You want to add a shaded region.  

* **Solution**  
```{r}
library(gcookbook) # For the data set

p <- ggplot(subset(climate, Source=="Berkeley"), aes(x=Year, y=Anomaly10y)) +
     geom_line()

p + annotate("rect", xmin=1950, xmax=1980, ymin=-1, ymax=1, alpha=.1,
             fill="blue")

```    

* **Discussion**  

## 7.6. Highlighting an Item
* **Problem**  
  You want to change the color of an item to make it stand out.  

* **Solution**  
```{R}
pg <- PlantGrowth                 # Make a copy of the PlantGrowth data
pg$hl <- "no"                     # Set all to "no"
pg$hl[pg$group=="trt2"] <- "yes"  # If group is "trt2", set to "yes"

ggplot(pg, aes(x=group, y=weight, fill=hl)) + geom_boxplot() +
    scale_fill_manual(values=c("grey85", "#FFDDCC"), guide=FALSE)
```

* **Discussion**  
```{r}
ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot() +
    scale_fill_manual(values=c("grey85", "grey85", "#FFDDCC"), guide=FALSE)
```

* **See Also**  

## 7.7. Adding Error Bars
* **Problem**  
  You want to add error bars to a graph.  

* **Solution**  
```{R}
library(gcookbook) # For the data set
# Take a subset of the cabbage_exp data for this example
ce <- subset(cabbage_exp, Cultivar == "c39")

# With a bar graph
ggplot(ce, aes(x=Date, y=Weight)) +
    geom_bar(fill="white", colour="black") +
    geom_errorbar(aes(ymin=Weight-se, ymax=Weight+se), width=.2)

# With a line graph
ggplot(ce, aes(x=Date, y=Weight)) +
    geom_line(aes(group=1)) +
    geom_point(size=4) +
    geom_errorbar(aes(ymin=Weight-se, ymax=Weight+se), width=.2)

```

* **Discussion**  
```{R}
# ====================================
ce

 Cultivar Date Weight        sd  n         se
      c39  d16   3.18 0.9566144 10 0.30250803
      c39  d20   2.80 0.2788867 10 0.08819171
      c39  d21   2.74 0.9834181 10 0.31098410



# ====================================
cabbage_exp

 Cultivar Date Weight        sd  n         se
      c39  d16   3.18 0.9566144 10 0.30250803
      c39  d20   2.80 0.2788867 10 0.08819171
      c39  d21   2.74 0.9834181 10 0.31098410
      c52  d16   2.26 0.4452215 10 0.14079141
      c52  d20   3.11 0.7908505 10 0.25008887
      c52  d21   1.47 0.2110819 10 0.06674995



# ====================================
# Bad: dodge width not specified
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(position="dodge") +
    geom_errorbar(aes(ymin=Weight-se, ymax=Weight+se),
                  position="dodge", width=.2)

# Good: dodge width set to same as bar width (0.9)
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(position="dodge") +
    geom_errorbar(aes(ymin=Weight-se, ymax=Weight+se),
                  position=position_dodge(0.9), width=.2)


# ====================================
pd <- position_dodge(.3)  # Save the dodge spec because we use it repeatedly

ggplot(cabbage_exp, aes(x=Date, y=Weight, colour=Cultivar, group=Cultivar)) +
    geom_errorbar(aes(ymin=Weight-se, ymax=Weight+se),
                  width=.2, size=0.25, colour="black", position=pd) +
    geom_line(position=pd) +
    geom_point(position=pd, size=2.5)

# Thinner error bar lines with size=0.25, and larger points with size=2.5
```

* **See Also**  

## 7.8. Adding Annotations to Individual Facets
* **Problem**  
  You want to add annotations to each facet in a plot.  

* **Solution**  
```{R}
# The base plot
p <- ggplot(mpg, aes(x=displ, y=hwy)) + geom_point() + facet_grid(. ~ drv)

# A data frame with labels for each facet
f_labels <- data.frame(drv = c("4", "f", "r"), label = c("4wd", "Front", "Rear"))

p + geom_text(x=6, y=40, aes(label=label), data=f_labels)

# If you use annotate(), the label will appear in all facets
p + annotate("text", x=6, y=42, label="label text")
```

* **Discussion**  
```{R}
# ====================================
# This function returns a data frame with strings representing the regression
# equation, and the r^2 value
# These strings will be treated as R math expressions
lm_labels <- function(dat) {
  mod <- lm(hwy ~ displ, data=dat)
  formula <- sprintf("italic(y) == %.2f %+.2f * italic(x)",
                     round(coef(mod)[1], 2), round(coef(mod)[2], 2))

  r <- cor(dat$displ, dat$hwy)
  r2 <- sprintf("italic(R^2) == %.2f", r^2)
  data.frame(formula=formula, r2=r2, stringsAsFactors=FALSE)
}

library(plyr) # For the ddply() function
labels <- ddply(mpg, "drv", lm_labels)
labels

 drv                              formula                  r2
   4 italic(y) == 30.68 -2.88 * italic(x) italic(R^2) == 0.65
   f italic(y) == 37.38 -3.60 * italic(x) italic(R^2) == 0.36
   r italic(y) == 25.78 -0.92 * italic(x) italic(R^2) == 0.04

# Plot with formula and R^2 values
p + geom_smooth(method=lm, se=FALSE) +
    geom_text(x=3, y=40, aes(label=formula), data=labels, parse=TRUE, hjust=0) +
    geom_text(x=3, y=35, aes(label=r2), data=labels, parse=TRUE, hjust=0)


# ====================================
# Find r^2 values for each group
labels <- ddply(mpg, "drv", summarise, r2 = cor(displ, hwy)^2)
labels$r2 <- sprintf("italic(R^2) == %.2f", labels$r2)
```

* **See Also**  
