---
title: "Chapter 8. Axes"
output: html_notebook
---

# Chapter 8. Axes

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [Chapter 8. Axes](#chapter-8-axes)
	- [8.1. Swapping X- and Y-Axes](#81-swapping-x-and-y-axes)
	- [8.2. Setting the Range of a Continuous Axis](#82-setting-the-range-of-a-continuous-axis)
	- [8.3. Reversing a Continuous Axis](#83-reversing-a-continuous-axis)
	- [8.4. Changing the Order of Items on a Categorical Axis](#84-changing-the-order-of-items-on-a-categorical-axis)
	- [8.5. Setting the Scaling Ratio of the X- and Y-Axes](#85-setting-the-scaling-ratio-of-the-x-and-y-axes)
	- [8.6. Setting the Positions of Tick Marks](#86-setting-the-positions-of-tick-marks)
	- [8.7. Removing Tick Marks and Labels](#87-removing-tick-marks-and-labels)
	- [8.8. Changing the Text of Tick Labels](#88-changing-the-text-of-tick-labels)
	- [8.9. Changing the Appearance of Tick Labels](#89-changing-the-appearance-of-tick-labels)
	- [8.10. Changing the Text of Axis Labels](#810-changing-the-text-of-axis-labels)
	- [8.11. Removing Axis Labels](#811-removing-axis-labels)
	- [8.12. Changing the Appearance of Axis Labels](#812-changing-the-appearance-of-axis-labels)
	- [8.13. Showing Lines Along the Axes](#813-showing-lines-along-the-axes)
	- [8.14. Using a Logarithmic Axis](#814-using-a-logarithmic-axis)
	- [8.15. Adding Ticks for a Logarithmic Axis](#815-adding-ticks-for-a-logarithmic-axis)
	- [8.16. Making a Circular Graph](#816-making-a-circular-graph)
	- [8.17. Using Dates on an Axis](#817-using-dates-on-an-axis)
	- [8.18. Using Relative Times on an Axis](#818-using-relative-times-on-an-axis)

<!-- tocstop -->

## 8.1. Swapping X- and Y-Axes
* **Problem**  
  You want to swap the x- and y-axes on a graph.  

* **Solution**  
```{R}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + coord_flip()
```

* **Discussion**  
```{R}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + coord_flip() +
    scale_x_discrete(limits=rev(levels(PlantGrowth$group)))
```

* **See Also**  

## 8.2. Setting the Range of a Continuous Axis
* **Problem**  
  You want to set the range (or limits) of an axis.  

* **Solution**  
```{r}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()
# Display the basic graph
p

p + ylim(0, max(PlantGrowth$weight))
```    
* **Discussion**  
```{R}
# ====================================
ylim(0, 10)
scale_y_continuous(limits=c(0, 10))


# ====================================
p + ylim(0, 10) + scale_y_continuous(breaks=NULL)

p + scale_y_continuous(breaks=NULL) + ylim(0, 10)


# ====================================
p + scale_y_continuous(limits=c(0, 10), breaks=NULL)


# ====================================
p + scale_y_continuous(limits = c(5, 6.5))  # Same as using ylim()

p + coord_cartesian(ylim = c(5, 6.5))


# ====================================
p + expand_limits(y=0)
```


## 8.3. Reversing a Continuous Axis
* **Problem**  
  You want to reverse the direction of a continuous axis.  

* **Solution**  
```{r}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + scale_y_reverse()

# Similar effect by specifying limits in reversed order
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() + ylim(6.5, 3.5)
```

* **Discussion**  
```{R}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
    scale_y_reverse(limits=c(8, 0))
```

* **See Also**  

## 8.4. Changing the Order of Items on a Categorical Axis
* **Problem**  
  You want to change the order of items on a categorical axis.  

* **Solution**  
```{R}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + scale_x_discrete(limits=c("trt1","ctrl","trt2"))
```

* **Discussion**  
```{R}
# ====================================
p + scale_x_discrete(limits=c("ctrl","trt1"))


# ====================================
p + scale_x_discrete(limits=rev(levels(PlantGrowth$group)))
```

* **See Also**  

## 8.5. Setting the Scaling Ratio of the X- and Y-Axes
* **Problem**  
  You want to set the ratio at which the x- and y-axes are scaled.  

* **Solution**  
```{r}
library(gcookbook) # For the data set

sp <- ggplot(marathon, aes(x=Half,y=Full)) + geom_point()

sp + coord_fixed()
```

* **Discussion**  
```{R}
# ====================================
sp + coord_fixed() +
    scale_y_continuous(breaks=seq(0, 420, 30)) +
    scale_x_continuous(breaks=seq(0, 420, 30))


# ====================================
sp + coord_fixed(ratio=1/2) +
    scale_y_continuous(breaks=seq(0, 420, 30)) +
    scale_x_continuous(breaks=seq(0, 420, 15))
```


## 8.6. Setting the Positions of Tick Marks
* **Problem**  
  You want to set where the tick marks appear on the axis.  

* **Solution**  
```{R}
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
    scale_y_continuous(breaks=c(4, 4.25, 4.5, 5, 6, 8))
```

* **Discussion**  
```{R}
# ====================================
seq(4, 7, by=.5)

4.0 4.5 5.0 5.5 6.0 6.5 7.0

5:10

5  6  7  8  9 10


# ====================================
# Set both breaks and labels for a discrete axis
ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
    scale_x_discrete(limits=c("trt2", "ctrl"), breaks="ctrl")

```    

* **See Also**  

## 8.7. Removing Tick Marks and Labels
* **Problem**  
  You want to remove tick marks and labels.  

* **Solution**  
```{r}
# ====================================
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + theme(axis.text.y = element_blank())


# ====================================
p + theme(axis.ticks = element_blank(), axis.text.y = element_blank())


# ====================================
p + scale_y_continuous(breaks=NULL)
```


* **Discussion**  

## 8.8. Changing the Text of Tick Labels
* **Problem**  
  You want to change the text of tick labels.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

hwp <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) +
          geom_point()

hwp

# ====================================
hwp + scale_y_continuous(breaks=c(50, 56, 60, 66, 72),
                         labels=c("Tiny", "Really\nshort", "Short",
                                  "Medium", "Tallish"))

```

* **Discussion**  
```{R}
# ====================================
footinch_formatter <- function(x) {
    foot <- floor(x/12)
    inch <- x %% 12
    return(paste(foot, "'", inch, "\"", sep=""))
}


# ====================================
footinch_formatter(56:64)

 "4'8\""  "4'9\""  "4'10\"" "4'11\"" "5'0\""  "5'1\""  "5'2\""  "5'3\""  "5'4\""



# ====================================
hwp + scale_y_continuous(labels=footinch_formatter)


# ====================================
hwp + scale_y_continuous(breaks=seq(48, 72, 4), labels=footinch_formatter)


# ====================================
timeHMS_formatter <- function(x) {
    h <- floor(x/60)
    m <- floor(x %% 60)
    s <- round(60*(x %% 1))                   # Round to nearest second
    lab <- sprintf("%02d:%02d:%02d", h, m, s) # Format the strings as HH:MM:SS
    lab <- gsub("^00:", "", lab)              # Remove leading 00: if present
    lab <- gsub("^0", "", lab)                # Remove leading 0 if present
    return(lab)
}


# ====================================
timeHMS_formatter(c(.33, 50, 51.25, 59.32, 60, 60.1, 130.23))

 "0:20"    "50:00"   "51:15"   "59:19"   "1:00:00" "1:00:06" "2:10:14"
```


## 8.9. Changing the Appearance of Tick Labels
* **Problem**  
  You want to change the appearance of tick labels.  

* **Solution**  
```{R}
# ====================================
bp <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot() +
      scale_x_discrete(breaks=c("ctrl", "trt1", "trt2"),
                       labels=c("Control", "Treatment 1", "Treatment 2"))
bp


# ====================================
bp + theme(axis.text.x = element_text(angle=90, hjust=1, vjust=.5))


# ====================================
bp + theme(axis.text.x = element_text(angle=30, hjust=1, vjust=1))
```

* **Discussion**  
```{r}
bp + theme(axis.text.x = element_text(family="Times", face="italic",
                                      colour="darkred", size=rel(0.9)))
```

* **See Also**  

## 8.10. Changing the Text of Axis Labels
* **Problem**  
  You want to change the text of axis labels.  

* **Solution**  
```{r}      
library(gcookbook) # For the data set

hwp <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
       geom_point()
# With default axis labels
hwp

# Set the axis labels
hwp + xlab("Age in years") + ylab("Height in inches")
```

* **Discussion**  
```{r}
# ====================================
hwp + labs(x = "Age in years", y = "Height in inches")

# ====================================
hwp + scale_x_continuous(name="Age in years")

# ====================================
hwp + scale_x_continuous(name="Age\n(years)")
```

## 8.11. Removing Axis Labels
* **Problem**  
  You want to remove the label on an axis.  

* **Solution**  
```{R}
p <- ggplot(PlantGrowth, aes(x=group, y=weight)) + geom_boxplot()

p + theme(axis.title.x=element_blank())
```

* **Discussion**  
```{R}
p + xlab("")
```

## 8.12. Changing the Appearance of Axis Labels
* **Problem**  
  You want to change the appearance of axis labels.  

* **Solution**  
```{r}
library(gcookbook) # For the data set

hwp <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

hwp + theme(axis.title.x=element_text(face="italic", colour="darkred", size=14))
```

* **Discussion**  
```{r}
hwp + ylab("Height\n(inches)") +
theme(axis.title.y=element_text(angle=0, face="italic", size=14))

hwp + ylab("Height\n(inches)") +
theme(axis.title.y=element_text(angle=90, face="italic", colour="darkred",
size=14))
```

* **See Also**  

## 8.13. Showing Lines Along the Axes
* **Problem**  
  You want to display lines along the x- and y-axes, but not on the other sides of the graph.  

* **Solution**  
```{r}
library(gcookbook) # For the data set

p <- ggplot(heightweight, aes(x=ageYear, y=heightIn)) + geom_point()

p + theme(axis.line = element_line(colour="black"))
```

* **Discussion**  
```{R}
# ====================================
p + theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(colour="black"))


# ====================================
# With thick lines, only half overlaps
p + theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(colour="black", size=4))

# Full overlap
p + theme_bw() +
    theme(panel.border = element_blank(),
          axis.line = element_line(colour="black", size=4, lineend="square"))
```

* **See Also**  

## 8.14. Using a Logarithmic Axis
* **Problem**  
  You want to use a logarithmic axis for a graph.  

* **Solution**  
```{r}
library(MASS) # For the data set

# The base plot
p <- ggplot(Animals, aes(x=body, y=brain, label=rownames(Animals))) +
     geom_text(size=3)
p

# With logarithmic x and y scales
p + scale_x_log10() + scale_y_log10()
```

* **Discussion**  
```{r}
# ====================================
Animals

                      body  brain
Mountain beaver      1.350    8.1
Cow                465.000  423.0
Grey wolf           36.330  119.5
...
Brachiosaurus    87000.000  154.5
Mole                 0.122    3.0
Pig                192.000  180.0



# ====================================
10^(0:3)

 1   10  100 1000



# ====================================
10^(-1:5)

 1e-01 1e+00 1e+01 1e+02 1e+03 1e+04 1e+05



# ====================================
p + scale_x_log10(breaks=10^(-1:5)) + scale_y_log10(breaks=10^(0:3))


# ====================================
library(scales)
p + scale_x_log10(breaks=10^(-1:5),
                  labels=trans_format("log10", math_format(10^.x))) +
    scale_y_log10(breaks=10^(0:3),
                  labels=trans_format("log10", math_format(10^.x)))


# ====================================
ggplot(Animals, aes(x=log10(body), y=log10(brain), label=rownames(Animals))) +
    geom_text(size=3)


# ====================================
library(scales)

# Use natural log on x, and log2 on y
p + scale_x_continuous(trans = log_trans(),
                       breaks = trans_breaks("log", function(x) exp(x)),
                       labels = trans_format("log", math_format(e^.x))) +
    scale_y_continuous(trans = log2_trans(),
                       breaks = trans_breaks("log2", function(x) 2^x),
                       labels = trans_format("log2", math_format(2^.x)))


# ====================================
library(gcookbook) # For the data set

ggplot(aapl, aes(x=date,y=adj_price)) + geom_line()

ggplot(aapl, aes(x=date,y=adj_price)) + geom_line() +
    scale_y_log10(breaks=c(2,10,50,250))
```


## 8.15. Adding Ticks for a Logarithmic Axis
* **Problem**  
  You want to add tick marks with diminishing spacing for a logarithmic axis.  

* **Solution**  
```{R}
library(MASS)   # For the data set
library(scales) # For the trans and format functions
ggplot(Animals, aes(x=body, y=brain, label=rownames(Animals))) +
    geom_text(size=3) +
    annotation_logticks() +
    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x))) +
    scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x)))
```                  

* **Discussion**  
```{R}
ggplot(Animals, aes(x=body, y=brain, label=rownames(Animals))) +
    geom_text(size=3) +
    annotation_logticks() +
    scale_x_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x)),
                  minor_breaks = log10(5) + -2:5) +
    scale_y_log10(breaks = trans_breaks("log10", function(x) 10^x),
                  labels = trans_format("log10", math_format(10^.x)),
                  minor_breaks = log10(5) + -1:3) +
    coord_fixed() +
    theme_bw()
```

* **See Also**  

## 8.16. Making a Circular Graph
* **Problem**  
  You want to make a circular graph.  

* **Solution**  
```{R}
# ====================================
library(gcookbook) # For the data set
wind

 TimeUTC  Temp WindAvg WindMax WindDir SpeedCat DirCat
       0  3.54    9.52   10.39      89    10-15     90
       5  3.52    9.10    9.90      92     5-10     90
      10  3.53    8.73    9.51      92     5-10     90
  ...
    2335  6.74   18.98   23.81     250      >20    255
    2340  6.62   17.68   22.05     252      >20    255
    2345  6.22   18.54   23.91     259      >20    255



# ====================================
ggplot(wind, aes(x=DirCat, fill=SpeedCat)) +
    geom_histogram(binwidth=15, origin=-7.5) +
    coord_polar() +
    scale_x_continuous(limits=c(0,360))
```

* **Discussion**  
```{R}
# ====================================
ggplot(wind, aes(x=DirCat, fill=SpeedCat)) +
    geom_histogram(binwidth=15, origin=-7.5, colour="black", size=.25) +
    guides(fill=guide_legend(reverse=TRUE)) +
    coord_polar() +
    scale_x_continuous(limits=c(0,360), breaks=seq(0, 360, by=45),
                       minor_breaks=seq(0, 360, by=15)) +
    scale_fill_brewer()


# ====================================
coord_polar(start=-45 * pi / 180)


# ====================================
# Put mdeaths time series data into a data frame
md <- data.frame(deaths = as.numeric(mdeaths),
                 month  = as.numeric(cycle(mdeaths)))

# Calculate average number of deaths in each month
library(plyr) # For the ddply() function
md <- ddply(md, "month", summarise, deaths = mean(deaths))
md

 month   deaths
     1 2129.833
     2 2081.333
 ...
    11 1377.667
    12 1796.500

# Make the base plot
p <- ggplot(md, aes(x=month, y=deaths)) + geom_line() +
     scale_x_continuous(breaks=1:12)

# With coord_polar
p + coord_polar()



# ====================================
# With coord_polar and y (r) limits going to zero
p + coord_polar() + ylim(0, max(md$deaths))


# ====================================
p + coord_polar() + ylim(0, max(md$deaths)) + xlim(0, 12)


# ====================================
# Connect the lines by adding a value for 0 that is the same as 12
mdx <- md[md$month==12, ]
mdx$month <- 0
mdnew <- rbind(mdx, md)

# Make the same plot as before, but with the new data, by using %+%
p %+% mdnew + coord_polar() + ylim(0, max(md$deaths))
```

* **See Also**  

## 8.17. Using Dates on an Axis
* **Problem**  
  You want to use dates on an axis.  

* **Solution**  
```{R}
# ====================================
# Look at the structure
str(economics)

'data.frame':   478 obs. of  6 variables:
 $ date    : Date, format: "1967-06-30" "1967-07-31" ...
 $ pce     : num  508 511 517 513 518 ...
 $ pop     : int  198712 198911 199113 199311 199498 199657 199808 199920 ...
 $ psavert : num  9.8 9.8 9 9.8 9.7 9.4 9 9.5 8.9 9.6 ...
 $ uempmed : num  4.5 4.7 4.6 4.9 4.7 4.8 5.1 4.5 4.1 4.6 ...
 $ unemploy: int  2944 2945 2958 3143 3066 3018 2878 3001 2877 2709 ...



# ====================================
ggplot(economics, aes(x=date, y=psavert)) + geom_line()

```

* **Discussion**  
```{r}
# ====================================
# Take a subset of economics
econ <- subset(economics, date >= as.Date("1992-05-01") &
                          date <  as.Date("1993-06-01"))

# Base plot - without specifying breaks
p <- ggplot(econ, aes(x=date, y=psavert)) + geom_line()
p


# ====================================
# Specify breaks as a Date vector
datebreaks <- seq(as.Date("1992-06-01"), as.Date("1993-06-01"), by="2 month")

# Use breaks, and rotate text labels
p + scale_x_date(breaks=datebreaks) +
    theme(axis.text.x = element_text(angle=30, hjust=1))


# ====================================
library(scales)
p + scale_x_date(breaks=datebreaks, labels=date_format("%Y %b")) +
    theme(axis.text.x = element_text(angle=30, hjust=1))


# ====================================
# Mac and Linux
Sys.setlocale("LC_TIME", "it_IT.UTF-8")

# Windows
Sys.setlocale("LC_TIME", "italian")
```

* **See Also**  

## 8.18. Using Relative Times on an Axis
* **Problem**  
  You want to use relative times on an axis.  

* **Solution**  
```{R}
# Convert WWWusage time-series object to data frame
www <- data.frame(minute = as.numeric(time(WWWusage)),
                  users  = as.numeric(WWWusage))

# Define a formatter function - converts time in minutes to a string
timeHM_formatter <- function(x) {
    h <- floor(x/60)
    m <- floor(x %% 60)
    lab <- sprintf("%d:%02d", h, m) # Format the strings as HH:MM
    return(lab)
}

# Default x axis
ggplot(www, aes(x=minute, y=users)) + geom_line()

# With formatted times
ggplot(www, aes(x=minute, y=users)) + geom_line() +
    scale_x_continuous(name="time", breaks=seq(0, 100, by=10),
                       labels=timeHM_formatter)
```

* **Discussion**  
```{R}
# ====================================
scale_x_continuous(breaks=c(0, 20, 40, 60, 80, 100),
    labels=c("0:00", "0:20", "0:40", "1:00", "1:20", "1:40"))


# ====================================
timeHM_formatter(c(0, 50, 51, 59, 60, 130, 604))

 "0:00" "0:50" "0:51" "0:59" "1:00" "2:10" "10:04"



# ====================================
timeHMS_formatter <- function(x) {
    h <- floor(x/3600)
    m <- floor((x/60) %% 60)
    s <- round(x %% 60)                       # Round to nearest second
    lab <- sprintf("%02d:%02d:%02d", h, m, s) # Format the strings as HH:MM:SS
    lab <- sub("^00:", "", lab)               # Remove leading 00: if present
    lab <- sub("^0", "", lab)                 # Remove leading 0 if present
    return(lab)
}


# ====================================
timeHMS_formatter(c(20, 3000, 3075, 3559.2, 3600, 3606, 7813.8))

 "0:20"    "50:00"   "51:15"   "59:19"   "1:00:00" "1:00:06" "2:10:14"
```

* **See Also**  
