---
title: "Chapter 10. Legends"
output: html_notebook
---

# Chapter 10. Legends

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [Chapter 10. Legends](#chapter-10-legends)
	- [10.1. Removing the Legend](#101-removing-the-legend)
	- [10.2. Changing the Position of a Legend](#102-changing-the-position-of-a-legend)
	- [10.3. Changing the Order of Items in a Legend](#103-changing-the-order-of-items-in-a-legend)
	- [10.4. Reversing the Order of Items in a Legend](#104-reversing-the-order-of-items-in-a-legend)
	- [10.5. Changing a Legend Title](#105-changing-a-legend-title)
	- [10.6. Changing the Appearance of a Legend Title](#106-changing-the-appearance-of-a-legend-title)
	- [10.7. Removing a Legend Title](#107-removing-a-legend-title)
	- [10.8. Changing the Labels in a Legend](#108-changing-the-labels-in-a-legend)
	- [10.9. Changing the Appearance of Legend Labels](#109-changing-the-appearance-of-legend-labels)
	- [10.10. Using Labels with Multiple Lines of Text](#1010-using-labels-with-multiple-lines-of-text)

<!-- tocstop -->

## 10.1. Removing the Legend
* **Problem**  
  You want to remove the legend from a graph.  

* **Solution**  
```{R}
# The base plot (with legend)
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()
p

# Remove the legend for fill
p + guides(fill=FALSE)

```    

* **Discussion**  
```{R}
# ====================================
# Remove the legend for fill
p + scale_fill_discrete(guide=FALSE)


# ====================================
p + theme(legend.position="none")

```   

## 10.2. Changing the Position of a Legend
* **Problem**  
  You want to move the legend from its default place on the right side.  

* **Solution**  
```{R}
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot() +
    scale_fill_brewer(palette="Pastel2")

p + theme(legend.position="top")
```

* **Discussion**  
```{R}
# ====================================
p + theme(legend.position=c(1,0), legend.justification=c(1,0))


# ====================================
p + theme(legend.position=c(1,1), legend.justification=c(1,1))


# ====================================
p + theme(legend.position=c(.85,.2)) +
    theme(legend.background=element_rect(fill="white", colour="black"))

# ====================================
p + theme(legend.position=c(.85,.2)) +
    theme(legend.background=element_blank()) +  # Remove overall border
    theme(legend.key=element_blank())           # Remove border around each item

```    

## 10.3. Changing the Order of Items in a Legend
* **Problem**  
  You want to change the order of the items in a legend.  

* **Solution**  
```{r}
# The base plot
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()
p

# Change the order of items
p + scale_fill_discrete(limits=c("trt1", "trt2", "ctrl"))
```

* **Discussion**  
```{R}
# ====================================
p + scale_fill_grey(start=.5, end=1, limits=c("trt1", "trt2", "ctrl"))


# ====================================
p + scale_fill_brewer(palette="Pastel2", limits=c("trt1", "trt2", "ctrl"))

```

* **See Also**  

## 10.4. Reversing the Order of Items in a Legend
* **Problem**  
  You want to reverse the order of items in a legend.  

* **Solution**  
```{R}
# The base plot
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()
p

# Reverse the legend order
p + guides(fill=guide_legend(reverse=TRUE))
```

* **Discussion**  
```{r}
scale_fill_hue(guide=guide_legend(reverse=TRUE))
```

## 10.5. Changing a Legend Title
* **Problem**  
  You want to change the text of a legend title.

* **Solution**  
```{R}
# The base plot
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()
p

# Set the legend title to "Condition"
p + labs(fill="Condition")

```

* **Discussion**  
```{R}
# ====================================
p + scale_fill_discrete(name="Condition")


# ====================================
library(gcookbook) # For the data set

# Make the base plot
hw <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
      geom_point(aes(size=weightLb)) + scale_size_continuous(range=c(1,4))

hw

# With new legend titles
hw + labs(colour="Male/Female", size="Weight\n(pounds)")


# ====================================
hw1 <- ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex, colour=sex)) +
       geom_point()

hw1


# ====================================
# Change just shape
hw1 + labs(shape="Male/Female")

# Change both shape and colour
hw1 + labs(shape="Male/Female", colour="Male/Female")


# ====================================
p + guides(fill=guide_legend(title="Condition"))
```


## 10.6. Changing the Appearance of a Legend Title
* **Problem**  
  You want to change the appearance of a legend title’s text.  

* **Solution**  
```{R}
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()

p + theme(legend.title=element_text(face="italic", family="Times", colour="red",
                                    size=14))
```

* **Discussion**  
```{R}
p + guides(fill=guide_legend(title.theme=
           element_text(face="italic", family="times", colour="red", size=14)))
```

* **See Also**  

## 10.7. Removing a Legend Title
* **Problem**  
  You want to remove a legend title.  

* **Solution**  
```{R}
ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot() +
    guides(fill=guide_legend(title=NULL))

```

* **Discussion**  
```{R}
scale_fill_hue(guide = guide_legend(title=NULL))
```

## 10.8. Changing the Labels in a Legend
* **Problem**  
  You want to change the text of labels in a legend.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

# The base plot
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()

# Change the legend labels
p + scale_fill_discrete(labels=c("Control", "Treatment 1", "Treatment 2"))


```

* **Discussion**  
```{R}
# ====================================
p + scale_fill_grey(start=.5, end=1,
                    labels=c("Control", "Treatment 1", "Treatment 2"))


# ====================================
p + scale_fill_discrete(limits=c("trt1", "trt2", "ctrl"),
                        labels=c("Treatment 1", "Treatment 2", "Control"))


# ====================================
# The base plot
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn, shape=sex, colour=sex)) +
     geom_point()
p

# Change the labels for one scale
p + scale_shape_discrete(labels=c("Female", "Male"))

# Change the labels for both scales
p + scale_shape_discrete(labels=c("Female", "Male")) +
    scale_colour_discrete(labels=c("Female", "Male"))
```

## 10.9. Changing the Appearance of Legend Labels
* **Problem**  
  You want to change the appearance of labels in a legend.  

* **Solution**  
```{R}
# The base plot
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()

# Change the legend label appearance
p + theme(legend.text=element_text(face="italic", family="Times", colour="red",
                                   size=14))
```

* **Discussion**  
```{R}
# Changes the legend title text for the fill legend
p + guides(fill=guide_legend(label.theme=
           element_text(face="italic", family="Times", colour="red", size=14)))
```
* **See Also**  

## 10.10. Using Labels with Multiple Lines of Text
* **Problem**  
  You want to use legend labels that have more than one line of text.  

* **Solution**  
```{R}
p <- ggplot(PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()

# Labels that have more than one line
p + scale_fill_discrete(labels=c("Control", "Type 1\ntreatment",
                                 "Type 2\ntreatment"))
```

* **Discussion**  
```{R}
library(grid)
p + scale_fill_discrete(labels=c("Control", "Type 1\ntreatment",
                                 "Type 2\ntreatment")) +
    theme(legend.text=element_text(lineheight=.8),
          legend.key.height=unit(1, "cm"))
```    
