---
title: "Chapter 12. Using Colors in Plots"
output: html_notebook
---

# Chapter 12. Using Colors in Plots

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [Chapter 12. Using Colors in Plots](#chapter-12-using-colors-in-plots)
	- [12.1. Setting the Colors of Objects](#121-setting-the-colors-of-objects)
	- [12.2. Mapping Variables to Colors](#122-mapping-variables-to-colors)
	- [12.3. Using a Different Palette for a Discrete Variable](#123-using-a-different-palette-for-a-discrete-variable)
	- [12.4. Using a Manually Defined Palette for a Discrete Variable](#124-using-a-manually-defined-palette-for-a-discrete-variable)
	- [12.5. Using a Colorblind-Friendly Palette](#125-using-a-colorblind-friendly-palette)
	- [12.6. Using a Manually Defined Palette for a Continuous Variable](#126-using-a-manually-defined-palette-for-a-continuous-variable)
	- [12.7. Coloring a Shaded Region Based on Value](#127-coloring-a-shaded-region-based-on-value)

<!-- tocstop -->

## 12.1. Setting the Colors of Objects
* **Problem**  
  You want to set the color of some geoms in your graph.  

* **Solution**  
```{R}
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point(colour="red")

library(MASS) # For the data set
ggplot(birthwt, aes(x=bwt)) + geom_histogram(fill="red"), colour="black"

```

* **Discussion**  

* **See Also**  

## 12.2. Mapping Variables to Colors
* **Problem**  
  You want to use a variable (column from a data frame) to control the color of geoms.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

# These both have the same effect
ggplot(cabbage_exp, aes(x=Date, y=Weight, fill=Cultivar)) +
    geom_bar(colour="black", position="dodge")

ggplot(cabbage_exp, aes(x=Date, y=Weight)) +
    geom_bar(aes(fill=Cultivar), colour="black", position="dodge")

# These both have the same effect
ggplot(mtcars, aes(x=wt, y=mpg, colour=cyl)) + geom_point()

ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point(aes(colour=cyl))

```

* **Discussion**  
```{R}
# ====================================
str(cabbage_exp)

'data.frame':   6 obs. of  6 variables:
 $ Cultivar: Factor w/ 2 levels "c39","c52": 1 1 1 2 2 2
 $ Date    : Factor w/ 3 levels "d16","d20","d21": 1 2 3 1 2 3
 $ Weight  : num  3.18 2.8 2.74 2.26 3.11 1.47
 $ sd      : num  0.957 0.279 0.983 0.445 0.791 ...
 $ n       : int  10 10 10 10 10 10
 $ se      : num  0.3025 0.0882 0.311 0.1408 0.2501 ...



# ====================================
# Convert to factor in call to ggplot()
ggplot(mtcars, aes(x=wt, y=mpg, colour=factor(cyl))) + geom_point()

# Another method: Convert to factor in the data
m <- mtcars              # Make a copy of mtcars
m$cyl <- factor(m$cyl)   # Convert cyl to a factor
ggplot(m, aes(x=wt, y=mpg, colour=cyl)) + geom_point()
```

* **See Also**  

## 12.3. Using a Different Palette for a Discrete Variable
* **Problem**  
  You want to use different colors for a discrete mapped variable.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

# Base plot
p <- ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) + geom_area()

# These three have the same effect
p
p + scale_fill_discrete()
p + scale_fill_hue()

# ColorBrewer palette
p + scale_fill_brewer()

```

* **Discussion**  
```{r}
# ====================================
# Basic scatter plot
h <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) +
    geom_point()

# Default lightness = 65
h

# Slightly darker
h + scale_colour_hue(l=45)


# ====================================
library(RColorBrewer)
display.brewer.all()


# ====================================
p + scale_fill_brewer(palette="Oranges")


# ====================================
p + scale_fill_grey()

# Reverse the direction and use a different range of greys
p + scale_fill_grey(start=0.7, end=0)
```

* **See Also**  

## 12.4. Using a Manually Defined Palette for a Discrete Variable
* **Problem**  
  You want to use different colors for a discrete mapped variable.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

# Base plot
h <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=sex)) + geom_point()

# Using color names
h + scale_colour_manual(values=c("red", "blue"))

# Using RGB values
h + scale_colour_manual(values=c("#CC6666", "#7777DD"))
```

* **Discussion**  
```{r}
# ====================================
levels(heightweight$sex)

 "f" "m"



# ====================================
h + scale_colour_manual(values=c(m="blue", f="red"))
```

* **See Also**  

## 12.5. Using a Colorblind-Friendly Palette
* **Problem**  
  You want to use colors that can be distinguished by colorblind viewers.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

# Base plot
p <- ggplot(uspopage, aes(x=Year, y=Thousands, fill=AgeGroup)) + geom_area()

# The palette with grey:
cb_palette <- c("#999999", "#E69F00", "#56B4E9", "#009E73", "#F0E442",
                "#0072B2", "#D55E00", "#CC79A7")

# Add it to the plot
p + scale_fill_manual(values=cb_palette)

# ====================================
c("#000000", "#E69F00", "#56B4E9", "#009E73", "#F0E442", "#0072B2", "#D55E00",
  "#CC79A7")

```

* **Discussion**  

* **See Also**  

## 12.6. Using a Manually Defined Palette for a Continuous Variable
* **Problem**  
  You want to use different colors for a continuous variable.  

* **Solution**  
```{R}
library(gcookbook) # For the data set

# Base plot
p <- ggplot(heightweight, aes(x=ageYear, y=heightIn, colour=weightLb)) +
     geom_point(size=3)

p

# With a gradient between two colors
p + scale_colour_gradient(low="black", high="white")

# A gradient with a white midpoint
library(scales)
p + scale_colour_gradient2(low=muted("red"), mid="white", high=muted("blue"),
    midpoint=110)

# A gradient of n colors
p + scale_colour_gradientn(colours = c("darkred", "orange", "yellow", "white"))


```

* **Discussion**  

* **See Also**  

## 12.7. Coloring a Shaded Region Based on Value
* **Problem**  
  You want to set the color of a shaded region based on the y value.  

* **Solution**  
```{R}
# ====================================
library(gcookbook) # For the data set

cb <- subset(climate, Source=="Berkeley")

cb$valence[cb$Anomaly10y >= 0] <- "pos"
cb$valence[cb$Anomaly10y < 0]  <- "neg"

cb

   Source Year Anomaly1y Anomaly5y Anomaly10y Unc10y valence
 Berkeley 1800        NA        NA     -0.435  0.505     neg
 Berkeley 1801        NA        NA     -0.453  0.493     neg
 Berkeley 1802        NA        NA     -0.460  0.486     neg
 ...
 Berkeley 2002        NA        NA      0.856  0.028     pos
 Berkeley 2003        NA        NA      0.869  0.028     pos
 Berkeley 2004        NA        NA      0.884  0.029     pos



# ====================================
ggplot(cb, aes(x=Year, y=Anomaly10y)) +
    geom_area(aes(fill=valence)) +
    geom_line() +
    geom_hline(yintercept=0)
```

* **Discussion**  
```{R}
# ====================================
# approx() returns a list with x and y vectors
interp <- approx(cb$Year, cb$Anomaly10y, n=1000)

# Put in a data frame and recalculate valence
cbi <- data.frame(Year=interp$x, Anomaly10y=interp$y)
cbi$valence[cbi$Anomaly10y >= 0] <- "pos"
cbi$valence[cbi$Anomaly10y < 0]  <- "neg"


# ====================================
ggplot(cbi, aes(x=Year, y=Anomaly10y)) +
    geom_area(aes(fill=valence), alpha = .4) +
    geom_line() +
    geom_hline(yintercept=0) +
    scale_fill_manual(values=c("#CCEEFF", "#FFDDDD"), guide=FALSE) +
    scale_x_continuous(expand=c(0, 0))

```    
