---
title: "Chapter 14. Output for Presentation"
output: html_notebook
---

# Chapter 14. Output for Presentation

<!-- toc orderedList:0 depthFrom:1 depthTo:6 -->

- [Chapter 14. Output for Presentation](#chapter-14-output-for-presentation)
	- [14.1. Outputting to PDF Vector Files](#141-outputting-to-pdf-vector-files)
	- [14.2. Outputting to SVG Vector Files](#142-outputting-to-svg-vector-files)
	- [14.3. Outputting to WMF Vector Files](#143-outputting-to-wmf-vector-files)
	- [14.4. Editing a Vector Output File](#144-editing-a-vector-output-file)
	- [14.5. Outputting to Bitmap (PNG/TIFF) Files](#145-outputting-to-bitmap-pngtiff-files)
	- [14.6. Using Fonts in PDF Files](#146-using-fonts-in-pdf-files)
	- [14.7. Using Fonts in Windows Bitmap or Screen Output](#147-using-fonts-in-windows-bitmap-or-screen-output)

<!-- tocstop -->

## 14.1. Outputting to PDF Vector Files
* **Problem**  
  You want to create a PDF of your plot.  

* **Solution**  
```{R}
# width and height are in inches
pdf("myplot.pdf", width=4, height=4)

# Make plots
plot(mtcars$wt, mtcars$mpg)
print(ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point())

dev.off()


# ====================================
# 8x8 cm
pdf("myplot.pdf", width=8/2.54, height=8/2.54)


# ====================================
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()

# Default is inches, but you can specify unit
ggsave("myplot.pdf", width=8, height=8, units="cm")
```

* **Discussion**  

* **See Also**  

## 14.2. Outputting to SVG Vector Files
* **Problem**  
  You want to create a scalable vector graphics (SVG) image of your plot.  

* **Solution**  
```{r}
svg("myplot.svg", width=4, height=4)
plot(...)
dev.off()

# With ggsave()
ggsave("myplot.svg", width=8, height=8, units="cm")
```

* **Discussion**  

## 14.3. Outputting to WMF Vector Files
* **Problem**  
  You want to create a Windows metafile (WMF) image of your plot.  

* **Solution**  
```{R}
win.metafile("myplot.wmf", width=4, height=4)
plot(...)
dev.off()

# With ggsave()
ggsave("myplot.wmf", width=8, height=8, units="cm")
```

* **Discussion**  

## 14.4. Editing a Vector Output File
* **Problem**  
  You want to open a vector output file for final editing.  

* **Solution**  

* **Discussion**  
```{r}
pdf("myplot.pdf", width=4, height=4, useDingbats=FALSE)

# or
ggsave("myplot.pdf", width=4, height=4, useDingbats=FALSE)
```

## 14.5. Outputting to Bitmap (PNG/TIFF) Files
* **Problem**  
  You want to create a bitmap of your plot, writing to a PNG file.  

* **Solution**  
```{R}
# ====================================
# width and height are in pixels
png("myplot.png", width=400, height=400)

# Make plot
plot(mtcars$wt, mtcars$mpg)

dev.off()


# ====================================
# width and height are in pixels
png("myplot-%d.png", width=400, height=400)

plot(mtcars$wt, mtcars$mpg)
print(ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point())

dev.off()


# ====================================
ppi <- 300
# Calculate the height and width (in pixels) for a 4x4-inch image at 300 ppi
png("myplot.png", width=4*ppi, height=4*ppi, res=ppi)
plot(mtcars$wt, mtcars$mpg)
dev.off()


# ====================================
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point()

# Default dimensions are in inches, but you can specify the unit
ggsave("myplot.png", width=8, height=8, unit="cm", dpi=300)
```

* **Discussion**  
```{R}
install.packages("Cairo")  # One-time installation
CairoPNG("myplot.png")
plot(...)
dev.off()
```

## 14.6. Using Fonts in PDF Files
* **Problem**  
  You want to use fonts other than the basic ones provided by R in a PDF file.  

* **Solution**  
```{R}
# ====================================
install.packages("extrafont")
library(extrafont)

# Find and save information about fonts installed on your system
font_import()

# List the fonts
fonts()


# ====================================
library(extrafont)
# Register the fonts with R
loadfonts()

# On Windows, you may need to tell it where Ghostscript is installed
# (adjust the path to match your installation of Ghostscript)
Sys.setenv(R_GSCMD = "C:/Program Files/gs/gs9.05/bin/gswin32c.exe")


# ====================================
library(ggplot2)
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
    ggtitle("Title text goes here") +
    theme(text = element_text(size = 16, family="Impact"))

ggsave("myplot.pdf", width=4, height=4)

embed_fonts("myplot.pdf")

```

* **Discussion**  

* **See Also**  

## 14.7. Using Fonts in Windows Bitmap or Screen Output
* **Problem**  
  You are using Windows and want to use fonts other than the basic ones provided by R for bitmap or screen output.  

* **Solution**  
```{R}
# ====================================
install.packages("extrafont")
library(extrafont)

# Find and save information about fonts installed on your system
font_import()

# List the fonts
fonts()


# ====================================
library(extrafont)
# Register the fonts for Windows
loadfonts("win")


# ====================================
library(ggplot2)
ggplot(mtcars, aes(x=wt, y=mpg)) + geom_point() +
    ggtitle("Title text goes here") +
    theme(text = element_text(size = 16, family="Georgia", face="italic"))

ggsave("myplot.png", width=4, height=4, dpi=300)
```

* **Discussion**  
